---
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---
```{r echo=FALSE, eval=TRUE}
library(TMB)
#precompile()
library(TMBhelper)
library(reshape2)
library(sp)
library(mgcv)
library(nlme)
library(Matrix)
library(plyr)
library(ggplot2)
library(rgeos)
#library(raster)
source("./src/utils.R")

#Compile and load c++ code-------
#code for linear+splines
compile("./src/spde_mixture.cpp")
dyn.load(dynlib("./src/spde_mixture"))

#code for linear only
compile("./src/spde_linear.cpp")
dyn.load(dynlib("./src/spde_linear"))

#code for splines only
compile("./src/spde_splines.cpp")
dyn.load(dynlib("./src/spde_splines"))

#--------------------------------
```


## **Mackerel Egg Production modelling using the Tweedie distribution and the SPDE approach in Template Model Builder** 

### Response to WGMEGS request
WGMEGS requested WGS2D to .... (to copy request from agenda)

#### Description of the situation
In this study we use a latent Gaussian Random field to model mackerel egg production (EP) in the spawning grounds of NEA mackerel. We assume that EP follows a Tweedie distribution. The code is set up to be able to handle models that include only linear, only splines, or finally both linear and smoothed terms. 

Summary of the data.

```{r,echo=FALSE, comment=""}
#Identify data source and check it is unique. Load it.
RData.fnames <- dir("./data",pattern="*.RData",full.names = TRUE)
if(length(RData.fnames)!=1) stop("Cannot find a single .RData file in ./data")
load(RData.fnames)

keep <- c("Year","Period","EP","depth","Temp","eastings","northings","psy4.npp")
dat <- dat[, colnames(dat) %in% keep]

if(any(is.na(dat))==T){print('NA values have been removed')
  dat <- dat[!rowSums(is.na(dat)) > 0,]
}


shapefile.fnames <- dir("./objects",pattern="*.RData",full.names = TRUE)
if(length(shapefile.fnames)!=1) stop("Cannot find a single .RData file in ./objects")
load(shapefile.fnames)

df <- dat 
coordinates(df) <- ~eastings+northings
df@proj4string <- ocean.refined@proj4string

rm(dat)
df$Year <- factor(df$Year)
```
* `EP` is the response variable
* `utm_x,utm_y`are the Universal Tranverse Mercator transformed coordinates (UTM32) in kilometers
* Remaining variables are covariates.

***
## Models tested

```{r,echo=TRUE}
#First choose the max edge
max_edge=90 #(in km)

#Create a list of candidate models

models <- list()
models$linear_noGMRF <- list()
models$linear_noGMRF$gam_setup = gam(EP ~ depth + Temp + psy4.npp
                                      ,data = df@data, fit=FALSE)

models$splines_noGMRF <- list()
models$splines_noGMRF$gam_setup = gam(EP ~
                                        s(depth, bs = "cs") +
                                        s(Temp, bs = "cs") +
                                        s(psy4.npp, bs = "cs"),
                                      data = df@data, fit=FALSE)

models$linear_GMRF <- list()
models$linear_GMRF$gam_setup = models$linear_noGMRF$gam_setup


models$splines_GMRF <- list()
models$splines_GMRF$gam_setup = models$splines_noGMRF$gam_setup
 

for(j in 1:length(models)) {
  source("./src/MEGS.R")
  models[[j]]$results$fitted <- models[[j]]$results$obj$report()$mu
  models[[j]]$results$AIC <- TMBAIC(models[[j]]$results$opt)
}
```

#Table of likelihoods
```{r,echo=FALSE}

AIC_df <- ldply(.data = models, function(x) {
  c(AIC=x$results$AIC,
    Likelihood=x$results$opt$objective,
    Time=x$results$timeUsed)
  })
names(AIC_df) <- c('Model type','AIC','Likelihood','Time used')
AIC_df
```

#Model results 
```{r,echo=FALSE}
best_model <- which(AIC_df[,'AIC']==min(AIC_df[,'AIC']))
#to bypass the automatic best model selection select the desired model manually by uncommenting the line below
#best_model <- 3
final_model <- models[[best_model]]
final_model$name <- AIC_df[best_model,1]
final_model$results$opt
final_model$results$rep
if (grepl('noGMRF',final_model$name)==TRUE) #this means the best model doesn't contain a GMRF
  { rangeIndex = NULL
    range = 'The model does not contain a GMRF'
    } else {
      rangeIndex = which(row.names(summary(final_model$results$rep,"report"))=="range")
      range = summary(final_model$results$rep,"report")[rangeIndex,]}
#fieldIndex = which(row.names(summary(rep,"report"))=="x")
print('Range parameter estimate is:')
range
```

#Fitted vs Observed
```{r,echo=FALSE}
#par(mfrow=c(1,1))
par(mfrow=c(2,nrow(AIC_df)/2))
for(j in 1:nrow(AIC_df)) {
plot(df$EP~models[[j]]$results$fitted, main=names(models)[[j]], ylab='observed', xlab='fitted'
     #, xlim=c(0,40), ylim=c(0,40)
     )
abline(a = 0, b = 1, col = "red")
}
```

#Plots of the GMRF
```{r,echo=FALSE}
if (grepl('noGMRF',final_model$name)==TRUE) #this means the best model doesn't contain a GMRF
  { range # i.e. print message that the model doesn't contain a GMRF
  } else {
    proj = inla.mesh.projector(mesh, dims = c(100,100), xlim=range(outer_boundary$loc[,1]), ylim=range(outer_boundary$loc[,2]))
    latentFieldMAP = final_model$results$rep$par.random[names(final_model$results$rep$par.random)=="x"] / exp(final_model$results$rep$par.fixed[which(names(final_model$results$rep$par.fixed)=="log_tau")])

par(mfrow=c(1,2))
image.plot(proj$x,proj$y, exp(inla.mesh.project(proj, latentFieldMAP)),col =  colorRampPalette(c("white","yellow", "red"))(12),
           xlab = 'Easting', ylab = 'Northing',
           main = "MAP estimate of the GMRF",
           cex.lab = 1.1,cex.axis = 1.1, cex.main=1, cex.sub= 1.1)
#contour(proj$x, proj$y,inla.mesh.project(proj, latentFieldMAP) ,add = T,labcex  = 1,cex = 1)
#plot(land_fine,add=T, col='grey')

latentFieldSD = sqrt(final_model$results$rep$diag.cov.random[names(final_model$results$rep$par.random)=="x"])/exp(final_model$results$rep$par.fixed[which(names(final_model$results$rep$par.fixed)=="log_tau")])
image.plot(proj$x,proj$y, exp(inla.mesh.project(proj, latentFieldSD)),col =  colorRampPalette(c("white","yellow", "red"))(12),
           xlab = 'Easting', ylab = 'Northing',
           main = "SD estimate of the GMRF",
           cex.lab = 1.1,cex.axis = 1.1, cex.main=1, cex.sub= 1.1)
#contour(proj$x, proj$y,inla.mesh.project(proj, latentFieldSD) ,add = T,labcex  = 1,cex = 1)
#plot(land_fine,add=T, col='grey')
      }
```

#Plot splines
```{r echo=FALSE}
if(grepl('splines|mixture', final_model$name)){
#muSpline <- sdSpline <- list()
muSpline = final_model$results$muSpline
#muSpline = muSpline[1:(length(muSpline)-(length(pred_df)*nrow(pred_df[[1]])))]
muSpline = split(muSpline, ceiling(seq_along(muSpline)/nrow(pred_df[[1]])))
sdSpline = final_model$results$sdSpline
#sdSpline = sdSpline[1:(length(sdSpline)-(length(pred_df)*nrow(pred_df[[1]])))]
sdSpline = split(sdSpline, ceiling(seq_along(sdSpline)/nrow(pred_df[[1]])))

par(mfrow=c(length(pred_df)/2,2))
for(i in 1:length(pred_df)){
  #i=1
  ylim.down = min(muSpline[[i]] - 1.96*sdSpline[[i]])
  ylim.up = max(muSpline[[i]] + 1.96*sdSpline[[i]])
plot(pred_df[[i]][,1], muSpline[[i]], lty=1,type = 'l', ylim = c(ylim.down,ylim.up),
     xlab = names(pred_df[[i]]), ylab = paste0("f(",names(pred_df[[i]]),")"),
     main = paste0("Spline for ", names(pred_df[[i]])))
lines(pred_df[[i]][,1], muSpline[[i]] - 1.96*sdSpline[[i]], lty=2)
lines(pred_df[[i]][,1], muSpline[[i]] + 1.96*sdSpline[[i]], lty=2)
abline(h = 0)
  }
}


```
