---
title: "Exploration of time series"
output:
  html_document: 
    df_print: paged
    keep_md: no
---

```{r setup, include="FALSE"}
#'========================================================================
# Time_series_exploration
#'========================================================================
#
# by Mark R Payne
# DTU-Aqua, Kgs. Lyngby, Denmark
# http://www.staff.dtu.dk/mpay
#
# Created Fri Jan 15 20:19:08 2021
#
# <Description>
#
# This work is subject to a Creative Commons "Attribution" "ShareALike" License.
# You are largely free to do what you like with it, so long as you "attribute"
# me for my contribution. See the fine print at the end for exact details.
#
# To do:
#
# Notes:
#   * Knitting of notebooks from Rstudio leads to all sorts of messy problems 
#     with the directory. Use here::here() to specify a location relative to
#     the project top directory
#
#'========================================================================

#Configure Knitr
knitr::opts_chunk$set(
  echo=FALSE,
	error = FALSE,
	message = FALSE,
	warning = FALSE,
	include = TRUE,
	results = "markup",
	fig.width = 10,
	fig.height=14/25*10,
  fig.align="center",
  fig.path=here::here("notebooks","figures","Time_series_exploration_"))

#Powerpoint needs 14 x 25cm figures

# knitr::opts_knit$set(
#    root.dir="..")

library(tidyverse)
library(lubridate)
library(here)
library(RSQLite)
source(here("src/00.Common_elements.r"))

theme_set(theme_bw(base_size = 14))
```

```{r}
#==========================================================================
# Configure ####
#==========================================================================

```

by Mark R Payne  <br>
DTU-Aqua, Kgs. Lyngby, Denmark  <br>
http://www.staff.dtu.dk/mpay  <br>
`r format(Sys.time(), "%a %b %d %Y, %H:%M:%S ")`

This notebook compares the outputs from the PredEng pipeline, in particularly to check for
similarities and differences between the different metrics that have been generated there.

```{r results="asis"}
#'========================================================================
# PCA analysis ####
#'========================================================================
#Import PCA databbase
PCA.clim <- readRDS(here("objects/PCA_clim.rds"))
PCA.ts   <- 
  readRDS(here("objects/PCA_full.rds")) %>%
  filter(PC=="PC1") %>%
  unite("resultName",PC,resultName) 

#Extract PCA metrics
PCA.import <- 
  PCA.clim %>%
  mutate(importance=map(PCA,~summary(.x)$importance[3,] %>% enframe(name="PC")))  %>%
  dplyr::select(-PCA,-raster) %>%
  unnest(importance) %>%
  mutate(PC=as.numeric(gsub("PC","",PC)))

PCA.loadings <-
  PCA.clim %>%
  mutate(loadings=map(PCA,~ summary(.x)$rotation[,1:3]),
         raster=map(raster,rasterToPoints),
         loadings.dat=map2(raster,loadings,~bind_cols(as_tibble(.x),as_tibble(.y)))) %>%
  dplyr::select(spName,resultName,loadings.dat) %>%
  unnest(loadings.dat) %>%
  dplyr::select(spName,resultName,x,y,PC1,PC2) %>%
  pivot_longer(starts_with("PC"),names_to = "PC",values_to = "loadings")

#'========================================================================
# Setup ####
#'========================================================================
#Import data
this.db <- dbConnect(SQLite(),here("data/PredEng/Blue_whiting_WGS2D_results.sqlite"))
stats <- 
  this.db %>%
  tbl("statistics") %>%
  collect() %>%
  filter(is.na(field),
         statName=="SDM") %>%
  #Merge in PCAs
  mutate(date=ymd(date)) %>%
  bind_rows(PCA.ts)%>%
  mutate(stat.type=str_sub(resultName,1,4))

#Calculate z scores
stats <-
  stats %>%
  group_by(resultName,spName) %>%
  mutate(value=scale(value)) %>%
  ungroup() 
  

#Get items to loop over
spNames <- 
  stats %>% 
  distinct(spName) %>%
  pull()


#Now loop
for(this.spName in spNames) {
  cat(sprintf("\n\n## %s \n \n",this.spName))
  
  #Plot PCA characteristics first
  p <- PCA.import %>%
    filter(spName == this.spName) %>%
    ggplot(aes(PC,value,colour=resultName,type=resultName)) +
    geom_point()+
    geom_line()+
    coord_cartesian(xlim=c(0,10)) 
  plot(p)
  
  
  p <- 
    PCA.loadings %>%
    filter(spName == this.spName) %>%
    ggplot(aes(x,y,fill=loadings))+
    annotation_map(map_data("world"),fill="black")+
    geom_raster()+
    facet_grid(PC ~resultName )+
    scale_fill_distiller(type="div")+
    coord_quickmap()+
    theme(panel.background = element_rect(fill = "grey80"),
          panel.grid = element_blank())
  
  plot(p)
  
  #PCA time series plot
  # PCA.ts %>%
  #   filter(spName==this.spName,
  #          year %in% climatology.yrs,
  #          month ==3,
  #          stat.type %in% c("PC1","PC2")) %>%
  #   mutate(lbl=sprintf("%02i",year %%100)) %>%
  #   pivot_wider(names_from=stat.type,values_from = value) %>%
  #   ggplot(aes(PC1,PC2,label=lbl))+
  #   geom_text()+
  #   facet_wrap(~resultName)

  #Time series plot
  p <- 
    stats %>%
    filter(spName == this.spName) %>%
    ggplot(aes(x=date,y=value,colour=resultName))+
    geom_line()+
    facet_wrap(~stat.type,ncol=1) 
  plot(p)
  
  #Correlation plot
  cor.dat <- 
    stats %>%
    filter(spName == this.spName) %>%
    dplyr::select(date,resultName,value) %>%
    pivot_wider(names_from=resultName,values_from = value) %>%
    dplyr::select(-date) %>%
    cor(use="pairwise.complete") %>%
    as_tibble(rownames="x") %>%
    pivot_longer(-x,names_to="y",values_to = "cor") %>%
    mutate(lbl=sprintf("%0.2f",round(cor,2))) %>%
    filter(x!=y)
  
  p <- 
    ggplot(cor.dat,aes(x,y))+
    geom_tile(aes(fill=cor))+
    geom_text(aes(label=lbl))+
    scale_fill_distiller(type="div",limits=c(-1,1))+
    theme(axis.text.x = element_text(angle=90,hjust=1))+
    labs(x="",y="") 
  plot(p)
  
}

```





***
*This work by Mark R Payne is licensed under a  Creative Commons
Attribution-NonCommercial-ShareAlike 3.0 Unported License. 
For details, see http://creativecommons.org/licenses/by-nc-sa/3.0/deed.en_US
Basically, this means that you are free to "share" and "remix" for 
non-commerical purposes as you see fit, so long as you "attribute" me for my
contribution. Derivatives can be distributed under the same or 
similar license.*

*This work comes with ABSOLUTELY NO WARRANTY or support.*

***



